Tree size, microhabitat, and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-09-18 manuscript/climate_sensitivity_draft.Rmd
more work on introduction

*First, what drives the observed tendency for large trees to suffer more during drought?*

@bennett_larger_2015 showed that in forests globally, large trees suffer greater growth reductions during drought, and numerous subsequent studies have reinforced this finding (Stovall et al. 2019, **REFS**). However, this analysis quantified tree size based on DBH, which has no direct mechanistic meaning. This study proposed two major mechanisms (besides insects) for the observed greater drougth growth reductions of large trees: (1) inherently greater biophysical challenge of being tall; (2) greater exposure of the crowns of large trees. Counteracting these effects, (3) the larger root systems of larger trees may confer an advantage in terms of allowing greater access to water, but it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure.

Canopy trees have lower drought resistance because they are exposed to higher solar radiation, greater wind speeds, and lower humidity. Alternatively, the generally supressed status of subcanopy trees may be insufficient to override the benefits of their buffered environment during drought.

@@

Here, we combine tree-ring records covering three droughts (1966, 1977, 1999), species functional and hydraulic trait measurments, and forest census data from a 25.6-ha ForestGEO plot in Virginia (USA) to test a series of hypotheses and associated specific predictions (Table 1) designed to yield functional understanding of how tree size, microenrivonment, and species' traits collectively shape drought responses. 
First, we focus on the role of tree size and its interaction with microenvironement. 
We confirm that, consistent with most forests globally [@bennett_larger_2015], larger-diameter trees have lower drought resistance in this forest, which is in an ecoregion represented by only one study in Bennett et al (2015) (H1.0). 
We then test hypotheses designed to disentangle the relative importance of tree height (H1.1), crown exposure (H1.2), and root water access, which should be greater for larger trees in dry but not perpetually wet microsites (H1.3). 
Second, we focus on the role of species' functional and hydraulic traits and their interaction with tree height. We hypothesize that drought resistance will follow predicted and observed patterns in relation to wood density, specific leaf area, and ring porosity (P2.1a-c), but that leaf hydraulic traits such as leaf area shrinkage upon dehydration and turgor loss point will prove better predictors (P2.1d-e). 
We then test whether these traits correlate with tree height (P2.2), potentially driving the observed tendency for taller trees to suffer more during drought (P2.3). 
Finally, we focused on variability among droughts, asking how  community resistance varied across droughts (H3.1) and whether the factors confirming vulnerability or resistance varied across droughts (H3.2).

>>>>

One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought? @bennett_larger_2015 showed that in forests globally, large trees suffer greater growth reductions during drought, and numerous subsequent studies have reinforced this finding (Stovall et al. 2019, **REFS**). However, this analysis quantified tree size based on DBH, which has no direct mechanistic meaning. This study proposed two major mechanisms--besides the tendency for bark beetles to preferentially attack larger trees (**REFS**)--for the observed greater drought growth reductions of large trees. First, [: (1)] inherently greater biophysical challenge of being tall]. Second, larger trees may have lower drought resistance because they are more often in the canopy, where they are exposed to higher solar radiation, greater wind speeds, and lower humidity (**REFS**). Alternatively, the generally supressed status of subcanopy trees may be insufficient to override the benefits of their buffered environment during drought. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems may confer an advantage in terms of allowing greater access to water (**REFS**); however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. Elucidating the mechanisms that place larger trees at a relative disadvantage during drought will be critical to better understanding and modeling drought impacts on forests.

@@

Here, we combine tree-ring records covering three droughts (1966, 1977, 1999), species functional and hydraulic trait measurments, and forest census data from a 25.6-ha ForestGEO plot in Virginia (USA) to test a series of hypotheses and associated specific predictions (Table 1) designed to yield functional understanding of how tree size, microenrivonment, and species' traits collectively shape drought responses. 
First, we focus on the role of tree size and its interaction with microenvironement. 
We confirm that, consistent with most forests globally [@bennett_larger_2015], larger-diameter trees have lower drought resistance in this forest, which is in an ecoregion represented by only one study in Bennett et al (2015) (H1.0). 
We then test hypotheses designed to disentangle the relative importance of tree height (H1.1), crown exposure (H1.2), and root water access, which should be greater for larger trees in dry but not in perpetually wet microsites (H1.3). 
Second, we focus on the role of species' functional and hydraulic traits and their interaction with tree height. We hypothesize that drought resistance will follow predicted and observed patterns in relation to wood density, specific leaf area, and ring porosity (P2.1a-c), but that leaf hydraulic traits such as leaf area shrinkage upon dehydration and turgor loss point will prove better predictors (P2.1d-e). 
We then test whether these traits correlate with tree height (P2.2), potentially driving the observed tendency for taller trees to suffer more during drought (P2.3). 
Finally, we focused on variability among droughts, asking how  community resistance varied across droughts (H3.1) and whether the factors confirming vulnerability or resistance varied across droughts (H3.2).
